Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-05-31 numerical-analysis.tex
Minor revision to analysis section.

Most analysis operations that can be performed on simulations that reside on
disk can be performed on in-memory simulations in \texttt{yt}.  This includes
projections (i.e., line integrals, both on- and off-axis), slices, 1-, 2-, and
3-D fluid phase distributions, calculation of derived quantities and arbitrary
data selection.  As of version 2.5 of \texttt{yt},  the Rockstar phase-space
halo finder \citep{2013ApJ...762..109B} can be executed through \texttt{yt} on
in-memory \enzo\ data.  Operations that currently cannot be conducted on
in-memory datasets are those that require spatial decomposition of data; for
instance, calculating marching cubes on a data object with \texttt{yt} is a
fully local operation and can be conducted \textit{in situ}.  However,
calculating topologically-connected sets requires a spatial decomposition of
data and thus cannot be conducted \textit{in situ}.  This prohibition extends
to most halo finding operations, most multi-level parallelism operations,
volume rendering and connected set identification.

@@

At user-specified intervals, the tracer particles output their data.  The data written out
typically includes the tracer particle's unique ID, position, velocity, density,
and temperature, but is easily extensible to output any grid-based
quantity that a user requires.  This capability has been used quite
effectively in several papers, including \citet{2010ApJ...715.1575S}
and \citet{2012ApJ...748...12S}. It is crucial to keep in mind that
tracer particles model a fixed number of Lagrangian fluid trajectories. The fluid
on the grid, however, models the motion of all the mass and represents
the average quantity in a grid cell's volume. Consequently, after a period of
evolution, tracer particles --even if they initially had the same
density distribution as the gas-- will not have the same density
distribution as the fluid. For example, they tend to accumulate at stagnation points
of the flow, and care has to be taken in using these particles
appropriately. Tracer particles are very useful in studies such as the variety of
histories of the hydrodynamic quantities in Lagrangian fluid elements
and when evaluating complex chemical and cooling models in regard to the
simpler ones used in the actual numerical evolution.

@@

\begin{eqnarray}
\nabla \cdot \vec{v} < 0 \\
\nabla T \cdot \nabla S > 0\\
T_2 > T_1 \\
\rho_2 > \rho_1,
\end{eqnarray}

where $\vec{v}$ is the velocity field, $T$ is the temperature, $\rho$ is the
density, and $S=T/\rho^{\gamma-1}$ is the entropy.  $T_2$ and $T_1$ are the post-shock~(downstream)
and pre-shock~(upstream) temperatures, respectively. An optional temperature floor
may be chosen, which is useful for situations such as cosmological simulations without a radiation
background.

>>>>

Most analysis operations that can be performed on data sets that reside on
disk can be performed on in-memory data sets in \texttt{yt}.  This includes
projections (i.e., line integrals, both on- and off-axis), slices, 1-, 2-, and
3-D fluid phase distributions, calculation of derived quantities and arbitrary
data selection.  As of version 2.5 of \texttt{yt},  the Rockstar phase-space
halo finder \citep{2013ApJ...762..109B} can be executed through \texttt{yt} on
in-memory \enzo\ data.  Operations that currently cannot be conducted on
in-memory datasets are those that require spatial decomposition of
data.  For 
instance, calculating marching cubes on a data object with \texttt{yt} is a
fully local operation and can be conducted \textit{in situ}.  However,
calculating topologically-connected sets requires a spatial decomposition of
data and thus cannot be conducted \textit{in situ}.  This prohibition extends
to halo finding operations other than Rockstar, most multi-level
parallelism operations, and volume rendering.

@@

\enzo saves tracer particle data at user-specified intervals,
independent of the intervals at which regular data sets are written.  The data written out
typically includes the tracer particle's unique ID, position, velocity, density,
and temperature, but is easily extensible to output any grid-based
quantity that a user requires.  This capability has been used quite
effectively in several papers, including \citet{2010ApJ...715.1575S}
and \citet{2012ApJ...748...12S}. It is crucial to keep in mind that
tracer particles model a fixed number of Lagrangian fluid trajectories. The fluid
on the grid, however, models the motion of all the mass and represents
the average quantity in a grid cell's volume. Consequently, after a period of
evolution, tracer particles --even if they initially had the same
density distribution as the gas-- will not have the same density
distribution as the fluid. For example, they tend to accumulate at stagnation points
of the flow, and care has to be taken in using these particles
appropriately. Tracer particles are very useful in studies such as the variety of
histories of the hydrodynamic quantities in Lagrangian fluid elements
and when evaluating complex chemical and cooling models in regard to the
simpler ones used in the actual numerical evolution.

@@

\begin{eqnarray}
\nabla \cdot \vec{v} < 0,\\
\nabla T \cdot \nabla S > 0,\\
T_2 > T_1,\\
\rho_2 > \rho_1,
\end{eqnarray}

where $\vec{v}$ is the velocity field, $T$ is the temperature, $\rho$ is the
density, and $S=T/\rho^{\gamma-1}$ is the entropy.  $T_2$ and $T_1$ are the post-shock~(downstream)
and pre-shock~(upstream) temperatures, respectively. An optional temperature floor
may be chosen, which is useful for situations such as cosmological simulations without a radiation
background where underdense gas in the intergalactic medium (IGM) can cool adiabatically to
very low temperatures.

### 2013-05-31 numerical-overview.tex
Minor revision to analysis section.

We have extended the \enzo\ code to include multiple models for star formation
and feedback.  It is difficult to directly simulate the formation of individual
stars in the context of simulations of galaxy formation and evolution due
to the immense computational cost.  Therefore, 
we adopt parametric methods which attempt
to model star forming regions.  One model is based on Kravtsov's method
 \citep{2003ApJ...590L...1K}, and the other is based on the method of 
\citet{CO1992}, which has been modified for use in an AMR code.  The basic
ideas behind the methods are straightforward and observationally motivated.
Similar (though somewhat more advanced) methods have been employed for the 
smoothed particle hydrodynamics
(SPH) method, most recently by Springel \& Hernquist 
\citep{sh03a,sh03b,hs03}. This 
method assumes a
multiphase IGM and has been shown to accurately reproduce the cosmic star
formation rate, and will be implemented into \enzo\ in the near future.

>>>>

We have extended the \enzo\ code to include multiple models for star formation
and feedback.  It is difficult to directly simulate the formation of individual
stars in the context of simulations of galaxy formation and evolution due
to the immense computational cost.  Therefore, 
we adopt parametric methods which attempt
to model star forming regions.  One model is based on Kravtsov's method
 \citep{2003ApJ...590L...1K}, and the other is based on the method of 
\citet{CO1992}, which has been modified for use in an AMR code.  The basic
ideas behind the methods are straightforward and observationally motivated.
Similar (though somewhat more advanced) methods have been employed for the 
smoothed particle hydrodynamics
(SPH) method, most recently by Springel \& Hernquist 
\citep{sh03a,sh03b,hs03}. This 
method assumes a
multiphase interstellar medium (ISM) and has been shown to accurately reproduce the cosmic star
formation rate, and will be implemented into \enzo\ in the near future.
